---
title: Basic oce object format
author: Clark Richards
date: 2016-05-21
---

`oce` uses S4 objects, with standard "slots" that separate:

* `metadata`: relevant metadata created from the input file (e.g. instrument parameters, cruise name, lead PI, etc)

* `data`: the data of the object (e.g. temperature, salinity, pressure, etc)

* `processingLog`: a log of processing steps, with times, applied to the object (e.g. creation, subsetting, etc)

Within an `oce` object, a particular item can be retrieved using the `[[` method (or accessor). E.g. to get the temperature from a ctd object:
```{r}
library(oce)
data(ctd)
T <- ctd[['temperature']]
```
This is the recommended way of accessing fields within an oce object, as it is less brittle to changes in the underlying object structure. 

The data can also be accessed directly, by following the path through the slots, e.g.
```{r}
slotNames(ctd)
str(ctd) # have a look at the "structure" of the object
T <- ctd@data$temperature
```
Note that the `@` character is used to designate slots, while `$` is used to designate the elements of each slot (which are themselves "lists"). This is an explicit way to access fields within the object. While the `[[` method is the recommended and preferred way of extracting fields from an object, there may be cases where it must be done directly (e.g. in re-assignment of an object to a field after passing through a function such as `lapply()`).

As demonstrated above, to get a look at the structure of the object, use the `str()` function (for "`str`ucture"):
```{r}
str(ctd)
```

